Flight clustering for route detection and CO{\Large{\textsubscript{2}}} emissions

### 2016-10-23 paper/paper.tex
paper

Most aircrafts have equipped with an ADB-S transmitter, with which they send
information about their status and the position continuously, to guarantee
better air traffic control. These messages are not encripted and, in most regions,
this technology will become mandatory in 2020. For these reasons OpenSky network
has been created \cite{opensky}.

OpenSky is an open source network that aggregates ADS-B messages collected by
volunteers around the globe. The dataset is accessible to everyone for
non-profit purposes and the data is provided in raw format, in order to
perform research and statistics about air traffic. \\
There are some limitation to the data that is collected by the network. Western
Europe is well, but only partially covered. In areas where coverage is spotty or
absent, it is not possible to detect complete airplanes informations.
Additionally, messages are collected by volunteers and there are no guarantees
about the reliability of the collected information.

@@

Routes are represented by an ordered list of the geographical sectors where the
aircrafts transit. Prior to devising the clustering algorithm, it is important to
define how to measure the distance between two different routes. Since each
sector is identified by a code, the string editing distance can be used to
measure the distance between two different routes. The distance is therefore
represented by the minimum number of operations -- insertion, deletion and
replacement -- to transform Route 1 into Route 2 or vice-versa. Different weights
can also be applied to different operations. The distance should not depend by
the length of the routes and for this reason it is divided by the sum of
the lengths of the routes. The final result represents the percentage of
difference between the routes\cite{routeclustering}.

@@

To answer these questions, we are going to use a dataset of raw ADS-B messages
encoded in AVRO format. This dataset provides information from the OpenSky
network for a full week of September 2016. 
The quantity of available messages is very high, but their content is missing
some important informations about the flights, such as airlines and airports. It
follows that there are other questions that must be faced in the development of
the project:

\begin{itemize}
  \item How to use raw ADS-B data to identify flights?
  \item How to use raw ADS-B to discover additional properties of flights, such
  as associated airline or CO2 consumption of the aircraft?
  \item How to identify the departure and arrival airports of a flight?
  \item How to identify standard routes from an unstructured group of flights?
  \item Is ADS-B data alone enough? Do we need external datasets?
\end{itemize}

@@

Analyzing the entire dataset requires a lot of computational power, hence the
need to use a cluster to perform analysis on the data. However, before executing
on the cluster, it is essential to be able to try the algorithms locally on a
small subset of data, in order to test their correctness end efficiency, and to
obtain useful statistics. For this reason, we used different technologies
depending on where we ran the analysis.

On the cluster, we chose to use \textit{Apache Spark} because it is extremely
fast while providing an easy and convenient development interface. In
particular, the \textit{Scala} version was chosen instead of PySpark because of
the strongly typing and the functional programming approach. In addition to
this, the Java library \textit{libadsb} was used to decode the raw ADS-B
messages.

@@

\begin{enumerate}
  \item Extract the raw messages and their timestamp from the avro-encoded data
  \item Decode the messages from the raw representation
  \item Filter out messages that are not position messages
  \item Order the messages according to their timestamp
  \item Group the messages by ICAO
  \item For each ICAO, go through the messages in order of timestamp and decode
  the positions
\end{enumerate}

Notice that, even if we were able to distribute the work by ICAO, there is still
a very high number of messages to be processed by a single worker node. In order
to avoid memory problems, we decided to split the decoding process between days,
therefore running the above algorithm separately for each day that we had
available in the dataset. This choice has obvious consequences on the phases
that follow, as we were not able to detect flights between two different days.
From our observations, however, we concluded that these flights are not very
frequent, so at the end of the day the final results did not suffer too much.

@@

Having extracted all position information from the raw avro files, there is now
the problem of identifying discrete flights from it.
Given a sequence of positions for a given aircraft, on a particular day, we
though about how to group this information in chunks, each chunk representing a
potential flight. An important insight is given by observing the data: ADS-B
messages systematically stop being received by the sensors when the aircraft
approaches the ground. It follows that there must be a pause in the data between
two different flights, and in particular between the landing of a flight and the
take-off of another.

@@

Position points for a given aircraft are thus grouped together according to a
simple clustering algorithm: all points are analyzed in order of timestamp, and
two subsequent points are considered belonging in the same cluster if and only
if they are no more than 20 minutes apart, their distance is no more than 20
kilometers and their altitudes differ no more than 200 meters.

@@

We acknowledge that the criteria according to which we group and filter discrete
flights are pretty restrictive, and that we inevitably throw away a lot of data
and possibly some flights in the process. However, given the unreliability of
the data and the difficult of our task, we wanted to deal with sufficiently
precise and detailed data, in order to make our life easier in the subsequent
phases of the processing. In particular, the restriction on the end points of a
flight being below 3000 m is the most important one for our purposes (as
explained below), and also the one that seems to filter out the majority of
potential flights.

@@

The route detection algorithm, as explained in details below, uses a
representation of flights and routes as a sequence of sectors, where each sector
represents portion of the world map. For this reason, the flights detection
phase outputs flights as sequences of sectors rather than position points. This
has two main advantages:

@@

We could have considered alternative ways to eliminate redundant information and
reduce the amount of data needed to represent a single flight, but the routes
detection algorithm would stil have transformed those representations to
sequences of sectors. It follows that converting directly from position points
to sectors was obviously the best approach.

In practice, a sequence of position points is translated in a sequence of
sectors by simply truncating the latitude and longitude part of a point to one
decimal digit. Strings of values with the same truncated latitude and
longitude represent the same sector, and are compressent into one. This process
yields sectors of about 10 km by 10 km, which is a good approximation for our
purposes, where sub-kilometer precision is not needed.

@@

The most important part of the process is the detection of routes between pairs
of airports. Standard routes between a pair airports are likely to be used by
the majority of flights between those airports. Our algorithm relies on the
assumption that the converse should also be true, namely that if a route is used
by the majority of flights going from an airport to another, than that route
must represent a (possibly, the) standard route between that pair of airports.

@@

\begin{enumerate}
  \item Apply the clustering algorithm to all the flights that have the
  current pair of airports as endpoints.
  \item Select the cluster that classifies the highest number of routes.
  \item Select a route that constitues the representative of
  that cluster as the one having the least distance between al the other
  routes in the cluster
  \item The representative of this cluster determines the standard route
  between the pair of airports
  \item All the routes classified by the other clusters represent routes that
  diverge from the standard route
\end{enumerate}

@@

\begin{itemize}
  \item The analysis clearly does not take into account flights, and therefore
  routes that cannot make through the route detection phase of the algorithm.
  Our dataset is not extremely reliable, so say the least. The OpenSky Network
  has, and the time or writing this paper, a fairly good coverage of western
  Europe, but some sports may still be poorly detected by the sensors. If a
  standard route happens to pass through an area that is poorly covered, it is
  likely that other, non standard routes will be selected by the algorithm as
  standard, just because they can be detected.
  \item There may be reasons or events, such as military conflicts, for which
  airplanes are forced to systematically deviate from a standard route, for a
  period of time that can span days or even months. These temporary anomalies
  clearly invalidate this assumption, and would go absolutely undetected by our
  algorithm, especially since it gives information over just a week of time.
\end{itemize}

These, however, are fundamental problems of the data we have available. They
affect the results of the algorithm, but do not depend on it so they cannot be
reduced using a more clever algorithm. We feel that, with the information
available and in particular without an external source
of standard routes, considering this assumption as true is the best
possible approach.

@@

\begin{itemize}
  \item There could be situations where clusters have the same amount of routes,
  or the numbers differ very slightly. In this cases, the accuracy of the
  results inevitably decreases.
  \item There could be areas that are poorly covered by the sensors, or that
  have few flights passing though it. A cluster with five routes would be
  selected against one with two or three, but clearly it does not have a strong
  argument in determining the standard route between two airports. There are two
  possible approaches to this. The first is to recognize that the informations
  available is not enough, and possibly leave some pairs of airports with
  undetermined standard routes. The other, and the one that we follow, is to
  just run the algorithm without taking into account how many routes are
  actually involved in the selection of the standard route. A possible
  refinement could be to add to the front-end visualization of the standard
  routes an indication of ``accuracy'' of a particular route based on the amount
  of data available in the detection.
\end{itemize}

@@

Our approach of clustering together routes represented as sequences of sectors
is inspired by the Leader Algorithm described in \cite{routeclustering}, but
differs from it in some ways. Sectors in \cite{routeclustering} are bounded
airspace regions under the control of a single air traffic controller or small
team, hence they cover a relatively large area. Flights with such subdivision
usually go through 5--10 sectors. Our sectors are smaller --- each flight has
approximately 20--40 of them --- therefore they represent a much more
fine-grained subdivision of the world map.

@@

A little research showed that an average aircraft procudes about 53 pounds of
carbon dioxide per mile. We used this value to compute the difference in CO2
production between the standard routes and the corresponding straight-line
route.

We also wanted to offer a visualization of the comsumption of fuel with respect
to a straight-line route. The visualization should show the points of a given
route on the map in different colors on a palette from greed to red, where hues
towards red would indicate a point of high consumption in the route.
To do this, we needed to attribute to each segment of the route a suitable value,
from which we could determine its color. Rather than trying to compute a value
of ``comsumption'' for each segment, we decided to determine it based on
how different the route is with respect to the straight-line route. The main
characteristic of a straight-line route is that the airplane always points to
the destination airport. Therefore, we considered as value of each segment the
angle between the direction that the aircraft is facing in that segment and the
direction of the destination airport. Higher values lead to more red segments.

@@

The position extraction phase was run on the entire dataset of raw AVRO-encoded
messages, and it was, and one could expect, the most time consuming. The entire
execution took about 48 hours. As a result of this phase, we were able to shrink
the data from about 85 GB per day to about 10 GB of position points per day.
This of course made the next phases of the algorithm more tractable. Position
data has then been processed by the flight detection algorithm, in about 24
hours. This phase, as we already explained, throws away a lot of data. Also,
flights are represented and saved to disk as sequences of sectors, where each
flight it composed of 20--50 sectors. This gives a very succint representation,
and we were able to store the output of this phase in just about 4 MB per day.
The small size of the flights data allowed us to run the route clustering
algorithm and all the additional JSON files generation for the visualization in
under 30 minutes. With such small execution time, we were able to test the
results of the algorithm several times, and improve its performance and results.

@@

This result should be attributed to the restrictive criteria of the
classification algorithm, as well as to the incomplete information provided by
the dataset. It should be recalled that, even though the OpenSky network
comprises at the moment of 70 active sensors, some spots in the western Europe
area may still be pooly covered, or not covered at all. Even a small poorly
covered are makes all information on the flights passing through it incomplete,
thus unsuitable for our analysis.

@@

In order to compute the amount of flights diverging from a standard route per
airline, we needed to associate a an airline to each flight. This is not a
completely trivial task, as airline information is not provided at all by the
data. Every aircraft, hence every flight, has an ICAO number associated to it,
from which it is possible to determine callsigns from the identification
messages in the dataset. In the case of commercial flights, the callsign of an
aircraft contains the three-letter code of the airline for which the
vehicle is flying for at the moment. Since aircraft are bought by airlines and
hardly transferred from an airline to another, it is safe to assume that only
one callsign is sufficient to determine the airline of an aircraft.
The mapping between three-letter codes of airlines and their name has been
done with the help of an external dataset.

@@

We were able to associate an airline to almost all of the flights the resulted
from the route clustering phase. From this information, we then determined how
many flights, per airline, follow and do not follow the standard route for their
particular pair of airports, as identified by our standard route detection.
We observed that the majority of flights do follow tha standard route for their
particular pair of departure and arrival airports. Table~\ref{airline-table}
shows an excerpt of the data that we were able to collect.

@@

@@

There is room for improvement in the route detection algorithm, for example
by tuning the way routes are clustered together, and in particular by selecting
a better distance function.
A possible alternative way to compute the distance could be to compute the area
delimited by two routes. This area, normalized according to the length of the
routes, could be a good estimate of how ``different'' two routes are.
A different approach could be to discard the current algorithm and instead use
an external dataset of standard routes. In this setting, standard routes need
not be computed. Flights that diverge from them can simply be discovered by
computing the distance from the standard route for the particular pair of
airports, and checking the result against a threshold.

Currently, the CO2 consumption for the flights is computed in a very simple way,
by multiplying the distance traveled by the airplanes by an estimate of the
quantity of CO2 that an average aircraft produces per mile. A better approach
could use the ICAO number of each single aircraft to index an external dataset
and discover its characteristics. These properties could be used to compute a
more precise and realistic estimate of CO2 production.

>>>>

Most aircrafts are equipped with an ADB-S transmitter, with which they send
information about their status and position continuously, to guarantee
better air traffic control. These messages are not encripted and, in most regions,
this technology will become mandatory in 2020. For these reasons OpenSky network
has been created \cite{opensky}.

OpenSky is an open source network that aggregates ADS-B messages collected by
volunteers around the globe. The dataset is accessible by everyone for
non-profit purposes and the data is provided in raw format, in order to
perform research and statistics about air traffic. \\
There are some limitation to the data that is collected by the network. Western
Europe is well, but only partially covered. In areas where coverage is spotty or
absent, it is not possible to detect complete airplane information.
Additionally, messages are collected by volunteers and there are no guarantees
about the reliability of the collected information.

@@

Routes are represented by an ordered list of the geographical sectors where the
aircrafts transit. Prior to devising the clustering algorithm, it is important to
define how to measure the distance between two different routes. Since each
sector is identified by a code, the string editing distance can be used to
measure the distance between two different routes. The distance is therefore
represented by the minimum number of operations -- insertion, deletion and
replacement -- to transform Route 1 into Route 2 or vice-versa. Different weights
can also be applied to different operations. The distance should not depend on
the length of the routes and for this reason it is divided by the sum of
the lengths of the routes. The final result represents the percentage of
difference between the routes\cite{routeclustering}.

@@

To answer these questions, we are going to use a dataset of raw ADS-B messages
encoded in AVRO format. This dataset provides information from the OpenSky
network for a full week of September 2016. The quantity of available messages is
very high, but their content is missing some important informations about the
flights, such as airlines and airports. It follows that there are other
questions that must be answered in the development of the project:

\begin{itemize}
  \item How to use raw ADS-B data to identify flights?
  \item How to use raw ADS-B data to discover additional properties of flights, such
  as associated airline or CO2 consumption of the aircraft?
  \item How to identify the departure and arrival airports of a flight?
  \item How to identify standard routes from an unstructured group of flights?
  \item Is ADS-B data alone enough to answer the above questions? Do we need
  external datasets?
\end{itemize}

@@

Analyzing the entire dataset requires a lot of computational power, hence the
need to use a cluster to perform the computations on the entire dataset.
However, before executing on the cluster, it is essential to be able to try the
algorithms locally on a small subset of the data, in order to test their correctness
end efficiency, and to obtain useful statistics.

On the cluster, we chose to use \textit{Apache Spark} because it is extremely
fast while providing an easy and convenient development interface. In
particular, the \textit{Scala} version was chosen instead of PySpark because of
the strong typing and the functional programming approach. In addition to
this, the Java library \textit{libadsb} was used to decode the raw ADS-B
messages.

@@

\begin{enumerate}
  \item Extract the raw messages and their timestamp from the avro-encoded data
  \item Decode the messages from the raw representation
  \item Filter out messages that are not position messages
  \item Group the messages by ICAO
  \item For each ICAO, order the messages according to their timestamp
  \item For each ICAO, go through the messages in order of timestamp and decode
  the positions
\end{enumerate}

Notice that, even if we were able to distribute the work by ICAO, there is still
a very high number of messages to be processed by a single worker node. In order
to avoid memory problems, we decided to split the decoding process between days,
therefore running the above algorithm separately for each day that we had
available in the dataset. This choice has obvious consequences on the phases
that follow, as we were not able to detect flights between two different days.
From our observations, however, we concluded that these flights are not very
frequent, so the final results did not suffer too much.

@@

Having extracted all position information from the raw AVRO files, there is now
the problem of identifying discrete flights from it.
Given a sequence of positions for a given aircraft, on a particular day, we
though about how to group this information in chunks, each chunk representing a
potential flight. An important insight is given by observing the data: ADS-B
messages systematically stop being received by the sensors when the aircraft
approaches the ground. It follows that there must be a pause in the data between
two different flights, and in particular between the landing of a flight and the
take-off of another.

@@

Position points for a given aircraft are thus grouped together according to a
simple clustering algorithm: all points are analyzed in order of timestamp, and
two subsequent points are considered belonging to the same cluster if and only
if they are no more than 20 minutes apart, their distance is no more than 20
kilometers and their altitudes differ no more than 200 meters.

@@

We acknowledge that the criteria according to which we group and filter discrete
flights are pretty restrictive, and that we inevitably throw away a lot of data
and possibly some flights in the process. However, given the unreliability of
the data and the difficulty of our task following from it, we wanted to deal
with sufficiently precise and detailed data, in order to make our life easier in
the next phases of the processing. In particular, the restriction on the end
points of a flight being below 3000 m is the most important one for our purposes
(as explained below), and unfortunately also the one that seems to filter out
the majority of potential flights.

@@

The route detection algorithm, as explained in details below, uses a
representation of flights and routes as a sequence of sectors, where each sector
represents a portion of the world map. For this reason, the flights detection
phase outputs flights as sequences of sectors rather than position points. This
has two main advantages:

@@

We could have considered alternative ways to eliminate redundant information and
reduce the amount of data needed to represent a single flight, but the routes
detection algorithm would still have transformed those representations to
sequences of sectors. It follows that converting directly from position points
to sectors was obviously the best approach.

In practice, a sequence of position points is translated in a sequence of
sectors by simply truncating the latitude and longitude part of a point to one
decimal digit. Contiguous sequences of values with the same truncated latitude and
longitude represent the same sector, and are compressent into one. This process
yields sectors of about 10 km by 10 km, which is a good approximation for our
purposes, where sub-kilometer precision is not needed.

@@

The most important part of the process is the detection of routes between pairs
of airports. Standard routes between a pair of airports are likely to be used by
the majority of flights between those airports. Our algorithm relies on the
assumption that the converse should also be true, namely that if a route is used
by the majority of flights going from an airport to another, than that route
must represent a (possibly, the) standard route between that pair of airports.

@@

\begin{enumerate}
  \item Apply the clustering algorithm to all the flights that have the
  current pair of airports as endpoints.
  \item Select the cluster that classifies the highest number of routes.
  \item Select a route that constitues the representative of
  that cluster as the one having the least distance between all the other
  routes in the cluster
  \item The representative of this cluster determines the standard route
  between the pair of airports
  \item All the routes classified by the other clusters represent routes that
  diverge from the standard route
\end{enumerate}

@@

\begin{itemize}
  \item The analysis clearly does not take into account flights, and therefore
  routes that cannot make through the route detection phase of the algorithm.
  Our dataset is not extremely reliable, so say the least. The OpenSky Network
  has, and the time or writing this paper, a fairly good coverage of western
  Europe, but some sports may still be poorly detected by the sensors. If a
  standard route happens to pass through an area that is poorly covered, it is
  likely that other, non standard routes will be selected by the algorithm as
  standard, just because they can be detected.
  \item There may be reasons or events, such as military conflicts, for which
  airplanes are forced to systematically deviate from a standard route, for a
  period of time that can span days or even months. These temporary anomalies
  clearly invalidate this assumption, and would go absolutely undetected by our
  algorithm. This is especially true in settings like the one we were working in,
  where information spans over a limited period of time.
\end{itemize}

These, however, are fundamental problems of the data we use. They
affect the results of the algorithm, but do not depend on it so they cannot be
reduced using a more clever algorithm. We feel that, with the information
available and in particular without an external source
of standard routes, considering the above assumption as true is the best
possible approach.

@@

\begin{itemize}
  \item There could be situations where clusters have the same amount of routes,
  or the numbers differ very slightly. In this cases, the accuracy of the
  results inevitably decreases.
  \item There could be areas that are poorly covered by the sensors, or that
  have few flights passing though it. A cluster with five routes would be
  selected against one with two or three, but clearly it does not have a strong
  argument in determining the standard route between two airports. There are two
  possible approaches to this. The first is to recognize that the information
  available is not enough, and possibly leave some pairs of airports with
  undetermined standard routes. The other, and the one that we follow, is to
  just run the algorithm without taking into account how many routes are
  actually involved in the selection of the standard route. A possible
  refinement could be to add to the front-end visualization of the standard
  routes an indication of ``accuracy'' of a particular route based on the amount
  of data available in the detection.
\end{itemize}

@@

Our approach of clustering together routes represented as sequences of sectors
is inspired by the Leader Algorithm described in \cite{routeclustering}, but
differs from it in some ways. Sectors in \cite{routeclustering} are bounded
airspace regions under the control of a single air traffic controller or small
team, hence they cover a relatively large area. Flights with such subdivision
usually go through 5--10 sectors. Our sectors are smaller --- each flight has
approximately 20--70 of them --- therefore they represent a much more
fine-grained subdivision of the world map.

@@

A little research showed that an average aircraft produces about 53 pounds of
carbon dioxide per mile. We used this value to compute the difference in CO2
production between the standard routes and the corresponding straight-line
route.

We also wanted to offer a visualization of the comsumption of fuel with respect
to a straight-line route. The visualization should show the points of a given
route on the map in different colors on a palette from greed to red, where hues
towards red would indicate a point of high consumption in the route.
To do this, we needed to attribute to each segment of the route a suitable value,
from which we could determine its color. Rather than trying to compute a value
of ``consumption'' for each segment, we decided to determine it based on
how different the route is with respect to the straight-line route. The main
characteristic of a straight-line route is that the airplane always points to
the destination airport. Therefore, we considered as value of each segment the
angle between the direction that the aircraft is facing in that segment and the
direction of the destination airport. Higher values lead to more red segments.

@@

The position extraction phase was run on the entire dataset of raw AVRO-encoded
messages, and it was, and one could expect, the most time consuming. The entire
execution took about 48 hours. As a result of this phase, we were able to shrink
the data from about 85 GB per day to about 10 GB of position points per day.
This of course made the next phases of the algorithm more tractable. Position
data has then been processed by the flight detection algorithm, in about 24
hours. This phase, as we already explained, throws away a lot of data. Also,
flights are represented and saved to disk as sequences of sectors, where each
flight it composed of 20--70 sectors. This gives a very succint representation,
and we were able to store the output of this phase in just about 4 MB per day.
The small size of the flights data allowed us to run the route clustering
algorithm and all the additional JSON files generation for the visualization in
under 30 minutes. With such small execution time, we were able to test the
results of the algorithm several times, and improve its performance and results.

@@

This result should be attributed to the restrictive criteria of the
classification algorithm, as well as to the incomplete information provided by
the dataset. It should be recalled that, even though the OpenSky network
comprises at the moment of 70 active sensors, some spots in the western Europe
area may still be pooly covered, or not covered at all. Even a small poorly
covered area makes all information on the flights passing through it incomplete,
thus unsuitable for our analysis.

@@

In order to compute the amount of flights diverging from a standard route per
airline, we needed to associate an airline to each flight. This is not a
completely trivial task, as airline information is not provided at all by the
data. Every aircraft, hence every flight, has an ICAO number associated to it,
from which it is possible to determine callsigns from the identification
messages in the dataset. In the case of commercial flights, the callsign of an
aircraft contains the three-letter code of the airline for which the
vehicle is flying for at the moment. Since aircraft are bought by airlines and
hardly transferred from an airline to another, it is safe to assume that only
one callsign is sufficient to determine the airline of an aircraft.
The mapping between three-letter codes of airlines and their name has been
done with the help of an external dataset.

@@

We were able to associate an airline to almost all of the flights that resulted
from the route clustering phase. From this information, we then determined how
many flights, per airline, follow and do not follow the standard route for their
particular pair of airports, as identified by our standard route detection.
We observed that the majority of flights do follow tha standard route for their
particular pair of departure and arrival airports. Table~\ref{airline-table}
shows an excerpt of the data that we were able to collect.

@@

As a final note, we were also able to conlude that ADS-B data alone is not
enough to extract information that is sufficiently relevant to other
applications. The large-scale data sets provided by OpenSky are extremely
useful, but additional sources of information are necessary if one wants to make
sense of their raw, unstructured data.

@@

There is room for improvement in the route detection algorithm, for example
by tuning the way routes are clustered together, and in particular by selecting
a better distance function.
A possible alternative way to estimate the distance could be to compute the area
delimited by two routes. This area, normalized according to the length of the
routes, could be a good indication of how ``different'' two routes are.
A different approach could be to discard the current algorithm and instead use
an external dataset of standard routes. In this setting, standard routes need
not be computed. Flights that diverge from them can simply be discovered by
computing the distance from the standard route for the particular pair of
airports, and checking the result against a threshold.

Currently, the CO2 consumption for the flights is determined in a very simple
way, by multiplying the distance traveled by the airplanes by an estimate of the
quantity of CO2 that an average aircraft produces per mile. A better approach
could use the ICAO number of each single aircraft to index an external dataset
containing its characteristics. These properties could be used to compute a
more precise and realistic estimate of CO2 production for each single aircraft.
